---
title: "Non-Partisan Elections come to Louisville"
format: html
---

NPR story: https://www.lpm.org/news/2024-04-03/what-you-need-to-know-about-louisvilles-likely-switch-to-nonpartisan-elections
2024 legislation: https://apps.legislature.ky.gov/record/24RS/hb388.html
Beshear's official veto: https://apps.legislature.ky.gov/record/24rs/hb388/veto.pdf 
- the mayor can't veto
- 

McCraney as independent former Democrat: https://www.courier-journal.com/story/news/politics/2025/01/29/louisville-metro-councilwoman-paula-mccraney-leaves-democratic-party/78020919007/

Making information harder to find is generally bad. The one advantage of polarized parties is that it gives you a clear vote.

(Maybe nod to proportional representation?, though I suppose for local needs we do want neighborhoods represented as neighborhoods)

Party of local control lmao 

I used the March 12, 2026 Council Meeting, which still has party affiliation listed 


There are currently 13 Democrats, 12 Republicans, and 1 Independent 

Load libraries
```{r}
library(tidyverse)
library(patchwork)
library(arcgis)
library(leaflet)
library(leaflet.extras2)
library(htmlwidgets)
```


```{r}
# url is listed at this website https://data.lojic.org/datasets/louisville-ky-metro-council-districts/explore?location=38.189800%2C-85.678250%2C10
data_url <- "https://gis.lojic.org/maps/rest/services/LojicSolutions/OpenDataPolitical/MapServer/8/query?outFields=*&where=1%3D1&f=geojson"

districts <- arc_read(data_url)

districts <- districts |>
    sf::st_transform('+proj=longlat +datum=WGS84')
```

```{r}
parties <- read_csv("council_party.csv")
```

```{r}
districts <- districts |>
    left_join(parties, join_by(COUNDIST == District))  
```

Note that Lojic has already complied and COUN_PARTY is NA for all districts

```{r}
partisan_map <- ggplot(data = districts) +
    geom_sf(aes(fill = Party), show.legend = FALSE) +
    #geom_sf_label(aes(label = COUNDIST)) +
    scale_fill_manual(values = c("D" = "blue", "R" = "red", "I"= "gray")) + 
    theme_void()

nonpartisan_map <- ggplot(data = districts) +
    geom_sf() +
    #geom_sf_label(aes(label = COUNDIST)) +
    theme_void()

transition_map <- partisan_map + nonpartisan_map +
   plot_annotation("Louisville's Metro Council Elections are becoming Nonpartisan", caption = "",  theme = theme(plot.title = element_text(face = "bold", size = 14, hjust = 0.5)))

transition_map
```


```{r}
custom_colors <- c("D" = "blue", "I"= "black", "R" = "red")
partypal <- colorFactor(palette = custom_colors, domain = districts$Party)

labels <- c(districts$COUNDIST, districts$COUN_NAME)

labels <- sprintf(
    "<strong>District %s</strong><br/>%s <br/>%s <br/>%s", districts$COUNDIST, districts$COUN_NAME, districts$COUN_EMAIL, districts$COUN_PHON) |>
        lapply(htmltools::HTML)

map <- leaflet(districts) |>
    addTiles() |>
    addPolygons(
        fillColor = ~partypal(Party),
        color = "black",
        weight = 2,
        fillOpacity = 0.3,
        label = labels,
        labelOptions = labelOptions(textsize = "12px")
    ) 
    
 map <- map |>
    addGeosearch(provider = geosearchProvider("OSM"),
        options = geosearchOptions(style = "bar"))


map
```

```{r}
saveWidget(map, file = "docs/council_districts.html", selfcontained = FALSE)
```

Getting the saved widget to work with ghost was not easy. I had to use an iframe with srcdoc

<iframe 
  srcdoc='YOUR_FULL_HTML_HERE'
  width="100%" 
  height="500px" 
  style="border: none;"
  loading="lazy"
  sandbox="allow-scripts">
</iframe>

Then I had to search and replaceall:

So your two substitutions, in this order:

& → &amp; (do this first, otherwise you'll double-escape the & in &#39;)
' → &#39;

Claude did somehow figure this out and it would not have happened from internet searches alone. 